# Working log {.unnumbered}

Here is the contents for Randy's thesis

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warnings = FALSE,
                      message = FALSE,
                      comment = "",
                      #results = "hide",
                      digits = 4,
                      error = FALSE)

## clean the R environment
# graphics.off()
# rm(list = ls())
# freshr::freshr()

## load packages
library(here, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(gtsummary, quietly = TRUE)
library(flextable, quietly = TRUE)

## check the directory for the file
# here::dr_here()
here::set_here()

## the figure or results should be saved 
# paste0("foldername/Sfilename_workingresult_", 
#      Sys.Date(), ".filetype")
```

```{r}
praise::praise()
```

## August 2024

### 2024-08-01 `r praise::praise()`

-   [x] committee meeting date before the evaluation date August 28th

-   [x] the proposal for the PhD thesis

-   [x] add the nonlinear simulation to the methods

-   [x] the paper plan see the Excel file

-   [x] the defense date

### Tables and Figures for the paper

- [x] Table1 for the EPIC data

- [x] the overall plots with a few simulations plots

- [x] flowchart for the algorithm

- [x] cross-validation for the time and matching number

- [x] the simulation results

- [x] against specification


