---
title: "Simulation25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warnings = FALSE,
                      message = FALSE,
                      comment = "#>",
                      #results = "hide",
                      digits = 4,
                      error = FALSE)

## clean the R environment
graphics.off()
rm(list = ls())
freshr::freshr()

## load packages
library(here, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(gtsummary, quietly = TRUE)
library(flextable, quietly = TRUE)
library(devtools, quietly = TRUE)

load_all()

## check the directory for the file
# here::dr_here()
here::set_here()

## the figure or results should be saved 
# paste0("foldername/Sfilename_workingresult_", 
#      Sys.Date(), ".filetype")
```

## Goal for this chapter

-   [x] simulation study for single and lmm 

  -   [x] matching number 10
  -   [x] matching number 25
  -   [x] matching number 30
  -   [x] matching number 50

-   [x] functions to pull out the results

-   [x] figures

    -   [x] bias
    -   [x] rmse
    -   [x] coverage

-   [x] You are unreal!

## pullout functions

need to be adjusted based on the simulation we run.

```{r}
#| eval: false
pullout <- function(location,
                    filename) {
  load(paste0(location, filename))
  sgl_n <- sgl_n50 %>% 
    map("centiles_observed") %>% 
    map_dfr(~mutate(., bias = abs(actual - `50`),
           mse = (actual - `50`)^2,
           cov50 = ifelse(actual <= `75` & actual >= `25`, 1, 0),
           cov80 = ifelse(actual <= `90` & actual >= `10`, 1, 0),
           cov90 = ifelse(actual <= `95` & actual >= `5`, 1, 0)) %>% 
             colMeans()) %>%
    dplyr::select(bias, mse, cov50, cov80, cov90) %>%
    colMeans()
    lmm <- list()
    lmm$bias <-  lmm_test %>%
      transmute(bias = abs(pred - ht)) %>%
      unlist() %>%
      mean()
    lmm$mse <-  lmm_test %>%
      transmute(mse = (pred - ht)^2) %>%
      unlist() %>%
      mean()
    lmm$cov50<- lmm_test %>%
      transmute(cov50 = ifelse(ht <= centile75 & ht >= centile25, 1, 0)) %>%
      unlist() %>%
      mean()
    lmm$cov80<- lmm_test %>%
      transmute(cov80  = ifelse(ht <= centile90 & ht >= centile10, 1, 0)) %>%
      unlist() %>%
      mean()
    lmm$cov90<- lmm_test %>%
      transmute(cov90 = ifelse(ht <= centile95 & ht >= centile05, 1, 0)) %>%
      unlist() %>%
      mean()

  return(list(sgl_n = sgl_n,
              lmm = lmm))
}

```

## LMM is better than SGL

the linear mixed model is doing better than single time prediction. as we can see in the first paper.

```{r}
#| eval: false
folder1 <- "single_t11_n50"
folder2 <- "results"

file_location <- paste0("results/", folder1, "/", folder2, "/")
files <- list.files(path = file_location, pattern = ".Rdata")

sim_ss <- map_dfr(files, ~pullout(file_location, .),
                     .progress = list(type = "iterator",
                                      format = "Calculating {cli::pb_bar} {cli::pb_percent}",
                                      clear = TRUE))
sim_ss <- sim_ss %>%
  mutate(term = rep(c("bias", "mse", "coverage50", 
                      "coverage80", "coverage90"), 
                    length(files)))

save(sim_ss, file = paste0("results/04_single_simulation_n50_", Sys.Date(), ".Rdata"))
```

```{r}
load("results/04_single_simulation_n30_2024-09-05.Rdata")
data <- sim_ss %>% 
  mutate(sgl_n = as.numeric(sgl_n),
         lmm = as.numeric(lmm)) %>%
  pivot_longer(cols = -term,
               names_to = "method",
               values_to = "values") %>% 
  mutate(method = factor(method, levels = c("sgl_n", "lmm"))) %>%
  filter(method %in% c("sgl_n",  "lmm"))
```

```{r}
library(latex2exp)
bias_plot <- data %>%
  filter(term == "bias") %>%
  ggplot(aes(x = method, y = as.numeric(values), group = factor(method))) +
  scale_x_discrete(labels = c("sgl_n" = parse(text = latex2exp::TeX("$S_{\\kappa = 10}$")),
                              "lmm" = "LMM")) +
  geom_point(aes(color = method)) +
  geom_boxplot(aes(fill = method)) +
  theme_bw() +
  labs(x = NULL) +
  labs(y = "MAE") +
  ggthemes::scale_fill_tableau("Jewel Bright") + 
  ## so far the best color composition
  ggthemes::scale_colour_tableau("Jewel Bright") +
  ## facet_wrap("group") +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.y = element_text(size = 16)) 
# bias_plot
```

```{r}
rmse_plot <- data %>%
  filter(term == "mse") %>% 
  mutate(values = sqrt(as.numeric(values))) %>%
  ggplot(aes(x = method, y = as.numeric(values), group = method)) +
  geom_point(aes(color = method)) +
  geom_boxplot(aes(fill = method)) +
  theme_bw() +
  labs(x = NULL) +
  labs(y = "RMSE") +
  ggthemes::scale_fill_tableau("Jewel Bright") + 
  ## so far the best color composition
  ggthemes::scale_colour_tableau("Jewel Bright") +
  ## facet_wrap("group") +
  theme(legend.position = "none") +
  scale_x_discrete(labels = c("sgl_n" = parse(text = latex2exp::TeX("$S_{\\kappa = 10}$")),
                              "lmm" = "LMM")) +
  theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.y = element_text(size = 16)) 
# rmse_plot
```

```{r}
cov50_plot <- data %>% 
  filter(term == "coverage50") %>% 
  ggplot(aes(x = method, y = as.numeric(values)),
         group = method) +
  geom_point(aes(color = method)) +
  geom_boxplot(aes(fill = method)) +
  theme_bw() +
  labs(x = NULL) +
  labs(y = "50% CR") +
  ggthemes::scale_fill_tableau("Jewel Bright") + 
  ## so far the best color composition
  ggthemes::scale_colour_tableau("Jewel Bright") +
  ## facet_wrap("group") +
  theme(legend.position = "none") +
  scale_x_discrete(labels = c("sgl_n" = parse(text = latex2exp::TeX("$S_{\\kappa = 10}$")),
                              "lmm" = "LMM")) +
  theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.y = element_text(size = 16)) 

## the tableau is in the ggthemes
# cov50_plot
```

```{r}
cov80_plot <- data %>% 
          filter(term == "coverage80") %>%
          ggplot(aes(x = method, y = as.numeric(values)),
                 group = method) +
          geom_point(aes(color = method)) +
          geom_boxplot(aes(fill = method)) +
          theme_bw() +
         labs(x = NULL) +
          labs(y = "80% CR") +
          ggthemes::scale_fill_tableau("Jewel Bright") + 
          ## so far the best color composition
          ggthemes::scale_colour_tableau("Jewel Bright") +
          ## facet_wrap("group") +
          theme(legend.position = "none") +
  scale_x_discrete(labels = c("sgl_n" = parse(text = latex2exp::TeX("$S_{\\kappa = 10}$")),
                              "lmm" = "LMM")) +
  theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.y = element_text(size = 16)) 
          ## the tableau is in the ggthemes
# cov80_plot
```

```{r}
cov90_plot <- data %>% 
          filter(term == "coverage90") %>%
          ggplot(aes(x = method, y = as.numeric(values)),
                 group = method) +
          geom_point(aes(color = method)) +
          geom_boxplot(aes(fill = method)) +
          theme_bw() +
          labs(x = NULL) +
          labs(y = "90% CR") +
          ggthemes::scale_fill_tableau("Jewel Bright") + 
          ## so far the best color composition
          ggthemes::scale_colour_tableau("Jewel Bright") +
          ## facet_wrap("group") +
          theme(legend.position = "none") +
  scale_x_discrete(labels = c("sgl_n" = parse(text = latex2exp::TeX("$S_{\\kappa = 10}$")),
                              "lmm" = "LMM")) +
  theme(axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.y = element_text(size = 16)) 
          ## the tableau is in the ggthemes
# cov90_plot
```

```{r fig.width=5, fig.height=3, warning=FALSE}
bias_plot
ggsave(paste0("figures/S04_t11_single_five_1bias_", Sys.Date(), ".png"))
rmse_plot
ggsave(paste0("figures/S04_t11_single_five_2rmse_", Sys.Date(), ".png"))
cov50_plot + 
  geom_hline(yintercept = 0.5, linetype="dashed", color = "indianred")
ggsave(paste0("figures/S04_t11_single_five_3cov50_", Sys.Date(), ".png"))
cov80_plot +
  geom_hline(yintercept = 0.8, linetype="dashed", color = "indianred")
ggsave(paste0("figures/S04_t11_single_five_4cov80_", Sys.Date(), ".png"))
cov90_plot +
  geom_hline(yintercept = 0.9, linetype="dashed", color = "indianred")
ggsave(paste0("figures/S04_t11_single_five_5cov90_", Sys.Date(), ".png"))


```

```{r, fig.height= 10, fig.width=6}
library(ggpubr)

figure1 <- ggarrange(bias_plot, 
                    # + theme(axis.text.x=element_blank()),
                    rmse_plot,
                    # + theme(axis.text.x=element_blank()),
                    # cov50_plot,
                    # + theme(axis.text.x=element_blank()), 
                    cov90_plot +
                      geom_hline(yintercept = 0.9,
                                 linetype = "dashed", 
                                 color = "indianred"),
                    # + theme(axis.text.x=element_blank()),
                    ncol = 1, nrow = 3)

figure1
ggsave(paste0("figures/S04_t11_single_five_6combine_", Sys.Date(), ".png"))
```
