---
title: "Practice time1 anchor4 p95"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warnings = FALSE,
                      message = FALSE,
                      comment = "#>",
                      #results = "hide",
                      digits = 4,
                      error = FALSE)

## clean the R environment
# graphics.off()
# rm(list = ls())
# freshr::freshr()

## load packages
library(here, quietly = TRUE)
library(lme4, quietly = TRUE)
library(devtools, quietly = TRUE)
library(optimx, quietly = TRUE)

devtools::load_all()
## check the directory for the file
# here::dr_here()
# here::set_here()

## the figure or results should be saved 
# paste0("foldername/Sfilename_workingresult_", 
#      Sys.Date(), ".filetype")
```


## Dynamic prediction Overall

```{r}
load("~/Desktop/plmphd/R/train_test.rda")
train1 <- train %>%
  group_by(id) %>%
  summarise(ht0 = min(ht),
            age0 = min(age))

train2 <- full_join(train, train1, by = "id") %>%
  mutate(sid = "S") %>%
  unite(id, sid, id, sep = "") %>% 
  mutate(sex = as.factor(sex))

test1 <- test %>%
  group_by(id) %>%
  summarise(ht0 = min(ht),
            age0 = min(age))

test2 <- full_join(test, test1, by = "id") %>%
  mutate(sid = "S") %>%
  unite(id, sid, id, sep = "") %>% 
  mutate(sex = as.factor(sex))
```


## Step1 Brokenstick

Here is the setups

```{r}
#| include: TRUE
bsk_knots <- c(3, 6, 9, 12, 14)
anchor <- c(4, 6, 8, 10, 12)
mhl_p <- 0.95

folder <- paste0("figures/bks", length(bsk_knots), "_anchor4_p095_time1/")
folder
```


```{r}
# ctl <- .makeCC("warning", tol = 1e-3)
bks_lmer1 <- lmer(ht ~ 1 + bs(time, knots = c(3, 6, 9, 12, 14), degree = 1) * sex + 
                   (1 + bs(time, knots = c(3, 6, 9, 12, 14), degree = 1) * sex| id),
                 # na.action = na.exclude,
                 control = lmerControl(check.nobs.vs.nRE = "warning",
                                         optCtrl = list(method = 'nlminb'),
                                         optimizer = "optimx"),
                 data = train2)

summary(bks_lmer1)
summary(rePCA(bks_lmer1))

save(bks_lmer1, file = paste0(folder, "P04_brokenstick_lmer1.Rdata"))
```

## Step2 dynamic prediction 

### Step2.1 Testing dynamic prediction

```{r}
test_baseline <- test2 %>%
  group_by(id) %>%
  mutate(group_index = row_number()) %>% 
  filter(time <= 1)

test_baseline %>% group_by(id) %>% summarize(n = n())

lp_test <- IndvPred_lmer(lmerObject = bks_lmer1,
                          data = train2,
                          newdata = test_baseline,
                          timeVar = "time",
                          outcomeVar = "ht",
                          idVar = "id",
                          lmer.fixed = "1 + bs(time, knots = c(3, 6, 9, 12, 14), degree = 1) * sex",
                          lmer.random = "1 + bs(time, knots = c(3, 6, 9, 12, 14), degree = 1) * sex",
                          M = 500,
                          times = anchor,
                          all_times = TRUE,
                          return_data = FALSE,
                          level = 0.9,
                          interval = "prediction",
                          seed = 555)
```


```{r}
#| eval: false
#| fig-width: 50
#| fig-height: 60

lp_test <- lp_test %>%
  as.data.frame() %>%
  mutate(ht = as.numeric(predicted_y),
         ht = round(ht, 2))
plot2 <- ggplot() +
  # geom_line(data = data1,
  #           aes(x = as.numeric(time), 
  #               y = ht, 
  #               group = id),
  #           color = "indianred") +
  # geom_point(data = data1,
  #            aes(x = as.numeric(time), 
  #               y = ht, 
  #               group = id)) +
  geom_line(data = lp_test,
            aes(x = as.numeric(time), 
                y = ht, 
                group = id),
            color = "darkgreen",
            alpha = 0.3) +
  geom_point(data = lp_test,
             aes(x = as.numeric(time), 
                y = ht, 
                group = id)) +
  geom_line(data = test2,
            aes(x = time,
                y = ht,
                group = id),
            color = "grey") +
  facet_wrap(~id, ncol = 30) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(title = "Individual Trajectories",
       x = "Time",
       y = "Outcome Score")

plot2

ggsave(paste0(folder, "P01_dynamic_predict_trajectories.png"),
       plot2,
       width = 45,
       height = 45,
       units = "in",
       dpi = 300)

```

### Step2.2 Training Tricky conditional prediction 

This part is very tricky.
Somehow I cannot directly predict the conditional prediction for the training data 
at the **anchor time sets**.

```{r}
#| eval: false
#| fig-wideth: 40
#| fig-height: 60

train3 <- train2 %>% 
  group_by(id) %>%
  slice(1L) %>% 
  dplyr::select(-time) %>%
  ungroup()

names(train2); names(train3)

# anchor <- eval(parse(text = anchor))
lb_train0 <- expand.grid(time = anchor,
                        id = train3$id)
lb_train1 <- lb_train0 %>% 
  as.data.frame() %>%
  full_join(train3, by = c("id"))

lb_train2 <- lb_train0 %>%  
  cbind(bs(lb_train0$time, 
           knots = c(3, 6, 9, 12), 
           degree = 1)) %>%
  as.data.frame()

lb_train3 <- lb_train2 %>% 
  full_join(train3, by = c("id")) %>% 
  mutate(`1:sexM` = `1` * ifelse(sex == "M", 1, 0),
         `2:sexM` = `2` * ifelse(sex == "M", 1, 0),
         `3:sexM` = `3` * ifelse(sex == "M", 1, 0),
         `4:sexM` = `4` * ifelse(sex == "M", 1, 0),
         `5:sexM` = `5` * ifelse(sex == "M", 1, 0)) %>%
  mutate(intercept = 1,
         sex = ifelse(sex == "M", 1, 0)) %>%
  dplyr::select(intercept, `1`:`5`, sex ,`1:sexM`:`5:sexM`) %>%
  mutate_all(as.numeric) %>%
  as.matrix()


re <- ranef(bks_lmer)[["id"]] %>% as.data.frame() %>%
  slice(rep(1:n(), each = 3))

dim(re) # 913 12
dim(lb_train) # 2739 12
A <- rowSums(lb_train * re)
lb_train1$pred <- predict(bks_lmer1,
                          ## marginal mean from the model
                          re.form = NA,
                          newdata = lb_train1) + A ## A is the random effect

## This code does not work
# lb_train$pred <- predict(bks_lmer1,
#                          re.form = NULL,
#                          ## re.form = "1 + bs(time, 
#                          ##            knots = c(3, 6, 9, 12, 14), 
#                          ##            degree = 1) * sex"
#                          newdata = lb_train)
lp_train <- train2 %>%
  mutate(pred = predict(bks_lmer1,
                        type = "response"))

plot2.5 <- ggplot() +
  geom_line(data = lp_train,
            aes(x = as.numeric(as.character(time)),
                y = ht,
                group = id),
            color = "black",
            alpha = 0.4) +
  geom_point(data = lp_train,
             aes(x = as.numeric(as.character(time)),
                y = ht,
                group = id),
             alpha = 0.5) +
  geom_point(data = lb_train,
            aes(x = as.numeric(as.character(time)), 
                y = pred, 
                group = id),
            color = "indianred") +
  facet_wrap(~id, ncol = 30) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(title = "Training Individual Trajectories",
       x = "Time",
       y = "Outcome Score")

plot2.5
ggsave(paste0(folder, "P01_brokenstick_dynamic_comparison.png"), 
       plot2.5, 
       width = 30,  
       height = 45, 
       units = "in", 
       dpi = 300) 
```

## Step3 Matching

This part is working surprisingly well.

When you think about the dynamic prediction! 
We only use observations before time2, extra pieces of B-spline cannot be
even conditional estimated for the prediction.

```{r}
#| eval: false

data2 <- lp_test %>%
  ungroup() %>%
  dplyr::select(id, time, ht) %>% 
  mutate(time = as.numeric(as.character(time)))

data1 <- lb_train1 %>% 
  ungroup() %>% 
  dplyr::select(id, time, ht = pred) %>% 
  mutate(time = as.numeric(as.character(time)))

subset <- data2 %>%
    group_by(id) %>%
    group_map(~ dis_match(lb_train = data1,
                          lb_test_ind = .,
                          train = train2,
                          match_methods = "euclidean",
                          id_var = id,
                          outcome_var = ht,
                          time_var = time,
                          match_number = 50,
                          match_time = FALSE,
                          match_plot = TRUE),
              .progress = TRUE)

plots <- map(subset, "plot")

library(gridExtra)
ggsave(paste0(folder, "P03_matching_subset_plot_eld50.png"), 
       marrangeGrob(grobs = plots,
                    ncol = 20, nrow = 23),
       width = 30,  
       height = 31, 
       units = "in", 
       dpi = 300)

```

## Step4 Final prediction

But the prediction is suprisingly well performed!

```{r}
#| eval: false
gf <- "ht ~ cs(time, df = 3)"
gs <- "~ cs(time, df = 1)"
results <- test2 %>%
  group_by(id) %>%
  group_map(~ as.data.frame(.)) %>% 
  map2(subset, 
       ~try(predict_gamlss(matching = .y$subset,
                           test_one = .x,
                           id_var = id,
                           time_var = time,
                           outcome_var = ht,
                           tmin = 0,
                           tmax = 20,
                           gamlss_formula = gf,
                           gamsigma_formula = gs,
                           predict_plot = TRUE) %>% 
              suppressMessages()),
       .progress = TRUE)

# subset %>% length() # 457
save(results, subset,
     file = paste0(folder, "P03_gamlss_subset_results_data_eld50.Rdata"))

#| eval: false
plots <- map(results, "predictive_centiles")

library(gridExtra)

ggsave(paste0(folder, "P03_final_predictive_plot_eld50.png"), 
       marrangeGrob(grobs = plots, ncol = 20, nrow = 25),
       width = 30, 
       height = 45, 
       dpi = 300)
```
