```{r setup, include=FALSE}
# knitr::opts_chunk(echo = TRUE)

## clean the R environment
rm(list = ls())
freshr::freshr()

## load packages
library(here)
library(tidyverse)
library(testthat)
library(usethis)
library(devtools)

load_all()


# ?testthat::test_file
# ?devtools::test
# path <- testthat_example("success")
# test_file(path)
# test_file(path, reporter = "minimal")

## check the directory for the file
here::dr_here()
here::set_here()
```

## initiation

```{r}
packageVersion("devtools")
## to create a package
# create_package("~/Desktop/project/plmlmm")
use_mit_license()

## look for R home directory
# R.home()
## look for R code base
# system.file()
## look at the files in the same directory
here::here() %>%
  list.files()
# list.dirs()
# here::here() %>%
#   list.dirs(full.names = FALSE,
#             recursive = TRUE)
```

## use_git

```{r}
## use git and github for the package
use_git()
usethis::use_git_remote("origin", 
                        url = NULL, 
                        overwrite = TRUE)

```

## use data

It’s a good idea to store the code that generates your individual internal data objects, as well as the use_data() call that writes all of them into R/sysdata.rda. This is workflow code that belongs below data-raw/, not below R/. usethis::use_data_raw() can be used to initiate the use of data-raw/ or to initiate a new .R script there. If your package is uncomfortably large, experiment with different values of compress in use_data(internal = TRUE).

```{r}
# usethis::use_data(tsa_train)
# usethis::use_data(tsa_test)
# usethis::use_data_raw("tsa_train")
# usethis::use_data_raw("tsa_test")

usethis::use_data(train, test,
                  internal = FALSE,
                  overwrite = TRUE)
```

## use package

```{r}
use_package("brokenstick", type = "Depends")
use_package("broom.mixed", type = "Depends")
use_package("plotly", type = "Depends")
use_package("gamlss", type = "Depends")
use_package("here", type = "Depends")
use_package("JMbayes", type = "Depends")
use_package("lme4", type = "Depends")
use_package("matrixcalc", type = "Depends")
use_package("MASS", type = "Depends")
use_package("nlme", type = "Depends")
use_package("tidyverse", type = "Depends")
use_package("tibble", type = "Depends")
use_package("readr")
use_package("rjags", type = "Depends")

use_package("stringr")
use_package("splines", type = "Depends")
use_package("shiny", type = "Depends")

devtools::install_github("Goodgolden/plm2024")
```

## use_r & use_test

```{r}
# use_r("00_utils")
# use_r("01_impute")
# use_r("02_distance")
# use_r("03_prediction")
# use_r("04_people-like-me")
# use_r("05_simulation")
# use_r("06_lmmpred)
use_r("07_summarization")
use_r("00_data")

use_test("01_impute")
# use_test("02_distance")
use_test("05_people-like-thee")
use_test("04_people-like-me")
```

## test_check

```{r}
test_check("plmphd")
```

## environment

```{r}
load_all()
exists("people_like_me",
       # where = globalenv(), 
       inherits = FALSE)
#> [1] FALSE

# ?exists Environment Access
env <- environment(plm2024)
exists("plm", 
       where = env,
       # where = globalenv(), 
       inherits = FALSE)
```

## documents

Use `Ctrl + .` in RStudio and start typing 'DESCRIPTION' to activate a helper that makes it easy to open a file for editing.

In addition to a filename, your hint can be a function name.

```{r}
## document the description
document()
## always check 
## it's good to establish the habit of checking this often.
check()
```

```{r}
load_all()
# rm(list = c("distance_df", "linear_impute", "match", "norm2", "not_all_na", "not_any_na",
#   "people_like_me", "plm_ind_plot", "predict_gamlss", "single_df", "soft_fun"))
## `covr` package to track 
## what proportion of your package’s source code
## is exercised by the tests.
# install.packages("covr")
library("covr")
report()
test()

# Q: what is poisson process
# A: a process in which events occur continuously and independently at a constant average rate.
# Q: what is the difference between poisson and exponential
# A: The Poisson distribution is the probability distribution of independent event occurrences in an interval. The exponential distribution is the probability distribution of the time between independent events.
# 

```

**Cmd + Shift + T (macOS)** for test

## readme

```{r}
## add description files to the package
# usethis::use_tidy_description()


## add datasets to the package


## build up a readme file and push into github
use_readme_rmd()
build_readme()

usethis::use_roxygen_md()
document()

```

Here is a typical sequence of calls when using devtools for package development:

-   Edit one or more files below R/.
-   document() (if you’ve made any changes that impact help files or NAMESPACE)
-   load_all()
-   Run some examples interactively.
-   test() (or test_file())
-   check()

```{r}
load_all()
plm2024:::plm_shiny()
```
