---
title: "02_miscellaneous"
author: "randy"
date: '`r Sys.Date()`'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warnings = FALSE,
                      message = FALSE,
                      comment = "#>",
                      #results = "hide",
                      digits = 4,
                      error = TRUE)

## clean the R environment
graphics.off()
rm(list = ls())
freshr::freshr()

## load packages
library(here, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(gtsummary, quietly = TRUE)
library(flextable, quietly = TRUE)

## check the directory for the file
# here::dr_here()
here::set_here()

## the figure or results should be saved 
# paste0("foldername/Sfilename_workingresult_", 
#      Sys.Date(), ".filetype")

```

# Notes

## Coding and functions

### Referencing variables

```{r}
data(mtcars)

## a problematic function with error

simple_function <- function(dataset, col_name){
  dataset %>%
    group_by(col_name) %>%
    summarise(mean_speed = mean(speed))
}
simple_function(cars, "dist")

## still get problem with the col_name
simple_function1 <- function(dataset, col_name){
  col_name <- as.name(col_name)
  dataset %>%
    group_by(col_name) %>%
    summarise(mean_speed = mean(speed))
}


simple_function1(cars, "dist")
```

To be able to do that, we need to use a framework that was introduced in the {tidyverse}, called tidy evaluation.

Most dplyr verbs use tidy evaluation in some way. Tidy evaluation is a special type of non-standard evaluation used throughout the tidyverse.

[What is data-masking and why do I need](https://rlang.r-lib.org/reference/topic-data-mask.html)

#### The `enquo()` - `!!()`

Now that col_name is (R programmers call it) quoted, or defused, we need to tell group_by() to evaluate the input as is. This is done with !!(), called the injection operator, which is another {rlang} function.

```{r}
simple_function2 <- function(dataset, col_name, value){
  col_name <- enquo(col_name)
  dataset %>%
    filter((!!col_name) == value) %>%
    summarise(mean_cyl = mean(cyl))
}

simple_function2(mtcars, am, 1)
```

enclosed `!!col_name` inside parentheses. This is because operators such as `==` have precedence over `!!`, so you have to be explicit. Also, notice that I didn’t have to quote 1. This is because it’s standard variable, not a column inside the dataset. Let’s make this function a bit more general.

```{r}
## take more than one col_name and use the value
simple_function3 <- function(dataset, filter_col, mean_col, value){
  filter_col <- enquo(filter_col)
  mean_col <- enquo(mean_col)
  dataset %>%
    filter((!!filter_col) == value) %>%
    summarise(mean((!!mean_col)))
}

simple_function3(mtcars, am, cyl, 1)
```

#### The `quos()` - `!!!()`

Because these dots contain more than one variable, you have to use `quos()` instead of `enquo()`. This will put the arguments provided via the dots in a list. Then, because we have a list of columns, we have to use `summarise_at()`.

So if you didn’t do them, go back to them and finish them first. Doing the exercise will also teach you what `vars()` and `funs()` are. The last thing you have to pay attention to is to use `!!!()` if you used `quos()`. So 3 `!` instead of only 2.

```{r}
## making the function more convienient
simple_function4 <- function(dataset, ...){
  col_vars <- quos(...)
  dataset %>%
    summarise_at(vars(!!!col_vars), 
                 list(~ mean(., trim = .2), ~ median(., na.rm = TRUE)))
}

simple_function4(mtcars, am, cyl, mpg)
```

Using ... with !!!() allows you to write very flexible functions.

If you need to be even more general, you can also provide the summary functions as arguments of your function, but you have to rewrite your function a little bit.

You might be wondering where the `quos()` went? Well because now we are passing two lists, a list of columns that we have to quote, and a list of functions, that we also have to quote, we need to use `quos()` when calling the function:

```{r}
simple_function5 <- function(dataset, cols, funcs){
  dataset %>%
    summarise_at(vars(!!!cols), funs(!!!funcs))
}

simple_function5(mtcars, quos(am, cyl, mpg), quos(mean, sd, sum))
```

To conclude this function, I should also talk about `as_label()` which allows you to change the name of a variable, for instance if you want to call the resulting column `mean_mpg` when you compute the mean of the `mpg` column:

```{r}
simple_function6 <- function(dataset, filter_col, mean_col, value){

  filter_col <- enquo(filter_col)
  mean_col <- enquo(mean_col)
  mean_name <- paste0("mean_", as_label(mean_col))
  
  dataset |>
    filter((!!filter_col) == value) |>
    summarise(!!(mean_name) := mean((!!mean_col)))
}

simple_function6(mtcars, filter_col = cyl,
                 value = 6, mean_col = mpg) 

```

Pay attention to the := operator in the last line. This is needed when using as_label().

#### Using `{{}}`

The previous section might have been a bit difficult to grasp, but there is a simplified way of doing it, which consists in using `{{}}`, introduced in `{rlang}` version 0.4.0. The suggested pronunciation of `{{}}` is curly-curly, but there is no consensus yet.

```{r}
how_many_na <- function(dataframe, column_name){
  dataframe %>%
    filter(is.na(column_name)) %>%
    count()
}

data(starwars)

how_many_na(starwars, hair_color)
```

```{r}
summarise_groups <- function(dataframe, grouping_var, column_name){

  grouping_var <- enquo(grouping_var)
  column_name <- enquo(column_name)
  mean_name <- paste0("mean_", as_label(column_name))

  dataframe %>%
    group_by(!!grouping_var) %>%  
    summarise(!!(mean_name) := mean(!!column_name, na.rm = TRUE))
}

summarise_groups1 <- function(dataframe, grouping_var, col_names) {
  dataframe %>%
    group_by({{ grouping_var }}) %>%
    summarise({{ col_names }} := mean({{ col_names }},
                                        na.rm = TRUE))
}
```

however, and if you want to modify the column names, for instance in this case return "mean_height" instead of height you have to keep using the `enquirer()`-`!!` syntax.

#### Anonymous functions

Have no identity, no name, but still do stuff! They will not live in the global environment. Like a person without a name, you would not be able to look the person up in the address book.

Since version 4.1, R introduced a short-hand for defining anonymous functions:

```{r}
map(c(1, 2, 3, 4), \(x)(1 / sqrt(x)))
```

## New pipe `|>`

1.  The command to clear all variables from the environment (workspace) `rm(list=ls())`
2.  In RStudio the keyboard shortcut for the pipe operator %\>% is Ctrl+Shift+M (Windows) or Cmd+Shift+M (Mac).
3.  In RStudio the keyboard shortcut for the assignment operator `<-` is Alt + - (Windows) or Option + - (Mac).
4.  In RStudio use Ctrl+L to clear all the code from your console.
5.  If you’re typing in a script in the source editor pane but you want to move the curser to the console use Ctrl+2. You can also use Ctrl+1 to move the curser back to the source editor.
6.  To run a line of code from the source editor use Ctrl+Enter (Windows) or Cmd+Enter (Mac).
7.  You can “tear” code panes or data view panes out of the RStudio window which can be particularly useful on big screens or when using multiple monitors.
8.  You can scroll through your command history by clicking Ctrl + ↑ (Windows) or Cmd + ↑ (Mac). 1. If you know what the line of code you’re looking for started with, type the first few characters and then press Ctrl/Cmd + ↑ and it will only search a matching subset of the history. 1.You can rename all instances of a variable name by highlighting one instance of the variable name and then using `Code > Rename` in Scope. This is better than using `Edit > Replace` and `Find` because it only looks for whole word matches.

## Example of in-paper cite Rmarkdown

@tbl-putting-functions-in-files presents some examples from the actual source of the [tidyr package](http://tidyr.tidyverse.org/) at version 1.1.2. There are some departures from the hard-and-fast rules given above, which illustrates that there's a lot of room for judgment here.

```{r include = FALSE}
library(tidyverse)

tidyr_separate <- tibble(
  `Organising principle` = "Main function plus helpers",
  `Source file` = "[tidyr/R/separate.R](https://github.com/tidyverse/tidyr/blob/v1.1.2/R/separate.R)",
  Comments = "Defines the user-facing `separate()` (an S3 generic), a `data.frame` method, and private helpers"
)

tidyr_rectangle <- tibble(
  `Organising principle` = "Family of functions",
  `Source file` = "[tidyr/R/rectangle.R](https://github.com/tidyverse/tidyr/blob/v1.1.2/R/rectangle.R)",
  Comments = "Defines a family of functions for \"rectangling\" nested lists (`hoist()` and the `unnest()` functions), all documented together in a big help topic, plus private helpers"
)

tidyr_uncount <- tibble(
  `Organising principle` = "One function",
  `Source file` = "[tidyr/R/uncount.R](https://github.com/tidyverse/tidyr/blob/v1.1.2/R/uncount.R)",
  Comments = "Defines exactly one function, `uncount()`, that's not particulary large, but doesn't fit naturally into any other `.R` file"
)
 
dat <- bind_rows(tidyr_uncount, tidyr_separate, tidyr_rectangle)
```

```{r}
#| echo: false
#| label: tbl-putting-functions-in-files
#| tbl-cap: Different ways to organize functions in files.
knitr::kable(dat)
```
