---
title: "People Like Me"
output: rmarkdown::html_vignette:
  css: style.css
vignette: >
  %\VignetteIndexEntry{People Like Me}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

set.seed(2024)
#Q: find a the color theme for the css file for html style template?
#A: https://www.w3schools.com/colors/colors_picker.asp
```

The goal of the people-like-me package is to provide a set of functions that can be used to predict the longitudinal trajectories for the new target individual. The package is designed to be user-friendly and easy to use. The package is still under development, but it is already functional and can be used to analyze data.

The difference between the people-like-me package and other linear mixed model based methods is that the people-like-me is designed based on curve matching and to predict a new out-of-sample target.
As we know, the prediction of the longitudinal trajectory is a challenging task, particularly when the target individual is not in the training dataset. The people-like-me package is designed to address this challenge by using the curve matching method, to find the matches who are most similar to the unknown target individual as a donor cohort, then to predict the longitudinal trajectory for the new target individual only based on the information from the donor cohort.

An alternative prediction based on linear mixed model is discussed by Dr. Dimitris Rizopoulos with [dynamic predictions](https://stats.stackexchange.com/questions/367521/how-does-a-fitted-linear-mixed-effects-model-predict-longitudinal-output-for-a-n). Here we summarize the key points for the linear mixed model based prediction: 

For example, ${\hat{y}}_{oj}$ denotes the observed outcome data for the new patient $j$, then you can first obtain an estimate, say ${\hat{b}}_j^{*}$ of his/her random effects from the posterior distribution $[b_j \mid {\hat{y}}_{oj}, \theta]$, where $\theta$ denotes the model parameters. For example, ${\hat{b}}_j^{*}$ is the mean of this posterior distribution. Given this estimate of his/her random effects, you calculate predictions using $x_j(t)\beta + z_j(t){\hat{b}}_j^{*}$, where $x_j(t)$ and $z_j(t)$ denote the design matrices for the fixed and random effects at the (future) time points of interest, and $\beta$ denotes the fixed effects. Standard errors for these predictions can be obtained using a Monte Carlo scheme.
For models fitted by `lme()` you can obtain these predictions using function `IndvPred_lme()` from package [JMbayes](https://cran.r-project.org/web/packages/JMbayes/index.html); also you can obtain the same type of individualized predictions using the `predict()` method for models fitted by the `mixed_model()` function of the [GLMMadaptive](https://drizopoulos.github.io/GLMMadaptive/) package.


The `plm` package is still under development, but it is already functional and can be used to analyze data.

```{r setup}
# library(plmphd)

# use_article("people-like-me")
```




